MLH1 Results Outline

### 2020-02-05 doc/Results.rmd
updated a chrm proportion figure

## MLH1 variance M1

@@

```{r M1.MLH1.var_cv, echo=FALSE}
#modeling within mouse variance
#use  DF.HetC.MixedModel.HQ

#HQ var.CO
MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
summary(MLH1var_M1)

@@

#CV
MLH1cv_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
summary(MLH1cv_M1)

@@

## MLH1 variance M2

$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$
M2 slash M3

```{r M2.MLH1.var, echo=FALSE}

@@

summary(MLH1var_M3)
anova(MLH1var_M3)  
as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))

@@

summary(MLH1cv_M3)

@@

summary(MLH1var_M3_Q12)
anova(MLH1var_M3_Q12)  
as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))

@@

summary(MLH1cv_M3_Q12)
anova(MLH1cv_M3_Q12)#sex large F

@@

## MLH1 variance M3

@@

```{r M3.MLH1.var, echo=FALSE}

@@

summary(MLH1var_M4)

@@

summary(MLH1cv_M4)
anova(MLH1cv_M4)  
as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))

@@

summary(MLH1var_M4_Q12)

@@

summary(MLH1cv_M4_Q12)
anova(MLH1cv_M4_Q12)

@@

#never make boxplot with mouse on X. mouse level boxplots should be faceted by subspecies 
#can't re-order the miceagg
#ff <- ggplot(MLH1_data[order(MLH1_data$strain), ], 
#             aes(x = as.factor(mouse), y = adj_nMLH1.foci, fill= strain ))+geom_boxplot()+ ggtitle("Boxplots of MLH1 distributions by mouse") + theme(axis.title.x=element_blank(),
#        axis.text.x=element_blank(),
#        axis.ticks.x=element_blank()) +scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", #"coral4","coral1", "#E69F00","red", "red4", "red3","purple","grey1", "grey", 
#        "grey","grey","grey",  "grey","grey","grey" )) +
# ylim( c((min(MLH1_data$adj_nMLH1.foci)-3) ), (max(MLH1_data$adj_nMLH1.foci)+5)) + facet_wrap(species~ sex, scales="free")
#don't show M.musculus by mouse
#scale_fill_manual(values=c("#56B4E9", "cadetblue", "lightblue", "coral4","coral1", "#E69F00","red", "red3", "red4","orange","yellow", "yellow3", 
 #       "green","green4","grey",  "grey","grey","grey" )) +

#order MLH1_data by strain!!
#colors_of_strains, colors .. 
#there is no order to 'mouse' so it looks shitty

@@

AP_strain.table.HQ <- ddply(MLH1_data, c("strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)
    )

@@

Fig1_v1_strain <- ggplot(AP_strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
  geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
  annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10,12, 14), #list of all x points 
           y = c(18.5,18.5,33,18.5, 18.5),

@@

> is there a strain effect in Dom?

@@

```{r male.show.plots, echo=FALSE}

@@

@@

booo <- summary(bhh11)
booo$coefficients

lm.male.strain <- lm(mean_co ~ strain, data= fomo.male_mouse.av)
ll <- summary(lm.male.strain)
ll$coefficients
#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

@@

MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)

@@

#ggplot

chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(~sex, scales = "free")
chrm.class.barplot

@@

#### HetC_M1

@@

#### HetC_M2

@@

#### HetC_M3

@@

$$mouse \ average \ SC  \ length ~=~ rand(strain) + \varepsilon $$

$$mouse \ average \ IFD \ trait  ~=~ rand(strain) + \varepsilon $$
The IFD traits are:

@@

# IFD

@@

@@

exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects

@@

@@

## Male Polymorphism Prediction

logistic regression
$$  Rec \ Group ~=~ mouse \ average \ IFD $$
For these tests I am looking at the single IFD measure from 2CO bivalents.

```{r male.IFD_strain, echo=FALSE}

# General CO Positions

@@

# Normalized 1CO positions

@@

@@

Males have much clearer separation of siscoten across chrm classes. This is emphasized when SC length is also plotted. It seems like musc males have higher amounts of this metric compared to Dom males (which might)

@@

### Male Polymorphism Prediction

@@

```{r male.Nrm1CO_strain, echo=FALSE}

@@

>>>>

### MLH1 variance M1

@@

```{r M1.MLH1.var_cv, echo=FALSE, include=FALSE}
#modeling within mouse variance
#use  DF.HetC.MixedModel.HQ

#HQ var.CO
MLH1var_M1 <- lmer( var ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
#summary(MLH1var_M1)

@@

#CV
MLH1cv_M1 <- lmer( cV ~ subsp * sex +(1|strain), data=DF.HetC.MixedModel.HQ )
#summary(MLH1cv_M1)

@@

### MLH1 variance M2

$$mouse CO metric  ~=~ subsp * sex * strain + \varepsilon$$

```{r M2.MLH1.var, echo=FALSE, include=FALSE}

@@

#summary(MLH1var_M3)
anova(MLH1var_M3)  
as.data.frame( round(summary(MLH1var_M3)$coefficients, 6))

@@

#summary(MLH1cv_M3)

@@

#summary(MLH1var_M3_Q12)
anova(MLH1var_M3_Q12)  
as.data.frame( round(summary(MLH1var_M3_Q12)$coefficients, 6))

@@

#summary(MLH1cv_M3_Q12)
anova(MLH1cv_M3_Q12)#sex large F

@@

### MLH1 variance M3

@@

```{r M3.MLH1.var, echo=FALSE, include=FALSE}

@@

#summary(MLH1var_M4)

@@

#summary(MLH1cv_M4)

anova(MLH1cv_M4)  
as.data.frame( round(summary(MLH1cv_M4)$coefficients, 6))

@@

#summary(MLH1var_M4_Q12)

@@

#summary(MLH1cv_M4_Q12)

anova(MLH1cv_M4_Q12)

@@

@@

#this move this to the setup file

strain.table.HQ_Fig <- ddply(MLH1_data, c("strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)
    )

@@

#might need to nudge the X values

Fig1_v1_strain <- ggplot(strain.table.HQ_Fig, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+

    geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+

    theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male

   #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
  annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10,12, 14), #list of all x points 
           y = c(18.5,18.5,33,18.5, 18.5),

@@

> Is there a strain effect in Dom?

@@

```{r male.show.plots, echo=FALSE, fig.width = 4, fig.height=2.5}
#try customize the plot size

@@

(Try removing the alpha legend)

@@

booo <- summary(bhh11)

lm.male.strain <- lm(mean_co ~ strain, data= fomo.male_mouse.av)
ll <- summary(lm.male.strain)

#even when all the other male strains are added -- the same result; PWD, MSM and SKIVE strain effects are significant.
```

@@

#no filename for add mouse, strain subsp
MELT_Chrm.Proportion <- add_category(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_sex(MELT_Chrm.Proportion)
MELT_Chrm.Proportion <- add_strain(MELT_Chrm.Proportion)

@@

#can't apply function use this
chrm.class.category_MELT$subsp <- ifelse(grepl("WSB", chrm.class.category_MELT$strain), "Dom", 
             ifelse(grepl("G", chrm.class.category_MELT$strain), "Dom",
                 ifelse(grepl("LEW", chrm.class.category_MELT$strain), "Dom", 
                ifelse(grepl("PERC", chrm.class.category_MELT$strain), "Dom",

          ifelse(grepl("CAST", chrm.class.category_MELT$strain), "Cast",
              ifelse(grepl("HMI", chrm.class.category_MELT$strain), "Cast",

                                       ifelse(grepl("MSM", chrm.class.category_MELT$strain), "Mol", #added Mol as distinct subsp class
                                      ifelse(grepl("MOLF", chrm.class.category_MELT$strain), "Mol",    
            ifelse(grepl("PWD", chrm.class.category_MELT$strain), "Musc", 
                     ifelse(grepl("CZECH", chrm.class.category_MELT$strain), "Musc", 
                 ifelse(grepl("AST", chrm.class.category_MELT$strain), "Musc",     
                      ifelse(grepl("TOM", chrm.class.category_MELT$strain), "Musc",  
                   ifelse(grepl("SKIVE", chrm.class.category_MELT$strain), "Musc",
                        ifelse(grepl("KAZ", chrm.class.category_MELT$strain), "Musc",

             ifelse(grepl("SPRET", chrm.class.category_MELT$strain), "Spretus",
               ifelse(grepl("SPIC", chrm.class.category_MELT$strain), "Spic",  "other"))))))))))))))))

chrm.class.category_MELT$subsp <- factor(chrm.class.category_MELT$subsp, ordered = TRUE,levels =c( "Dom",
                                                                 "Cast",
                                                                 "Musc",
                                                                 "Mol",
                                                                 "Spretus","Spic",
                                                                 "other") )
#ggplot

chrm.class.barplot <- ggplot(data = chrm.class.category_MELT, aes(y = mean.percent.type, x = strain, fill=as.factor(variable)) ) + geom_col(position = "dodge")+ggtitle("Chrm Proportions from BivData Data")+ylim(c(0,1)) + facet_wrap(sex~subsp, scales = "free")
chrm.class.barplot

@@

#### SC_HetC_M1

@@

#### SC_HetC_M2

@@

#### SC_HetC_M3

@@

M1 glm model for fixed strain effect across male averages.

$$mouse \ average \ SC  \ length ~=~ strain + \varepsilon $$
M2 anova for glm, of each subset, does at least one of the strains differ from others?

M3 logistic regression, can a bivalent metric predict the high or low group?

$$ log( strain ~=~ \ mouse \ average \ SC  \ length ) $$

The IFD traits are:

@@

# Q1. IFD

@@

### IFD_HetC_M1

@@

exactRLRT(lmer.IFD.RAW_MM.M1)#random strain not sig
as.data.frame(round(summary(lmer.IFD.RAW_MM.M1)$coefficients, 6))#these Dont return p values for the fixed effects
```

@@

### IFD_HetC_M2

```{r IFD_HetC_M2, echo=FALSE}

### IFD_HetC_M3

```{r IFD_HetC_M3, echo=FALSE}

@@

# Q1. General CO Positions

@@

# Q1. Normalized 1CO positions

**for all the test blocks below, integrate the 3 other metrics; siscoten, telomere, centromere distance.

@@

#### Nrm_1COpos_HetC_M1

#### Nrm_1COpos_HetC_M2

#### Nrm_1COpos_HetC_M3

@@

Males have much clearer separation of siscoten across chrm classes. This is emphasized when SC length is also plotted. It seems like musc males have higher amounts of this metric compared to Dom males.

@@

# Q2 SC

@@

```{r male.Nrm1CO_strain, echo=FALSE, eval=FALSE}

@@

## SC_male.poly_glm

remeber the 3 predictions for subsp;

i) Dom, no difference

ii) Musc, PWD and SKIVE greater than others

iii) Mol, MSM > MOLF

```{r Q2.SC.ploy.mal, echo=FALSE}
# 1 way anova, is at least 1 strain different from the rest?
# given a lm, how sig is the strain effect

#1. determine which strain effects are sig
#use mouse biv average
#Male.poly.Mouse.Table_BivData_4MM

glm1.SC.male.poly <- glm(mean_SC ~ subsp * strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm1.SC.male.poly) 
#SC many strains are different (more than the gwRR)
#subsps are sig

glm2.SC.male.poly <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM)
#summary(glm2.SC.male.poly)

#Dom
glm_SC.male_Dom <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Dom",])

#summary(glm_SC.male_Dom)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",])

#summary(glm_SC.male_Musc)

#Mol
glm_SC.male_Mol <- glm(mean_SC ~ strain, data= Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",])

#summary(glm_SC.male_Mol)

#Logistic regressions

log.SC.male.poly <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))

                #Curated_BivData[(Curated_BivData$sex == "male") & (Curated_BivData$subsp == "Musc"),],

#summary(log.SC.male.poly)

log.SC.male.poly_musc <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))

#summary(log.SC.male.poly_musc)
#not sig -- when just looking at Musc

#significant in predicting group
log.SC.male.poly_mol <- glm(Rec.group ~ mean_SC, 
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))

#summary(log.SC.male.poly_mol)
#not sig -- when just looking at Mol
```

#### Q2 SC by class

I think these comparisons should be done with the **Bivalent observations**

The logistic regression showed rec groups could be predicted by SC length - re-run them while seperating out by chrm class

```{r Q2.SC.Chrm.Class, echo=FALSE}

#1CO
log.SC_male_1CO.poly <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_1CO.poly)

log.SC_male_1CO.Musc <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_1CO.Musc)
#double check these directions

log.SC_male_1CO.Mol <- glm(Rec.group ~ mean.SC_1CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))

#summary(log.SC_male_1CO.Mol)

#2CO
log.SC_male_2CO.poly <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_2CO.poly)

log.SC_male_2CO.Musc <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_2CO.Musc)

log.SC_male_2CO.Mol <- glm(Rec.group ~ mean.SC_2CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_2CO.Mol)

#3CO
log.SC_male_3CO.poly <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM,  family=binomial(link="logit"))
#summary(log.SC_male_3CO.poly)

log.SC_male_3CO.Musc <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Musc",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Musc)

log.SC_male_3CO.Mol <- glm(Rec.group ~ mean.SC_3CO,
              data=Male.poly.Mouse.Table_BivData_4MM[Male.poly.Mouse.Table_BivData_4MM$subsp == "Mol",],  family=binomial(link="logit"))
#summary(log.SC_male_3CO.Mol)

These are the plots of the logistic regression.

```{r Q2.SC.log.regression.show.plots, echo=FALSE}

plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.blah.mouse.av

plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.logReg.mouse.av_1CO

plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.logReg.mouse.av_2CO

the plot of log regression for 1COs is wacky/reversed. Musc -- the mean SC lengths for 1COs are longer in the low group

> **Summary**

When the SC lengths are divided by co number / class, -- difference in direction of the logistic regression. I think the shorter 1CO bivalents in the high rec strain reflects the clustering of groups by SC length. (IE the scatter plots of SC length, binned by chrm class on the X).

Each point is a mouse average SC. When all bivalents are pooled, the logistic regression SC lengths predict the groups. When the data is divided by CO number, 1CO's have a higher SC length in the low group - in the low group more chrms, including the longer ones have 1CO - whereas in the high group, only shorter chrms (SC) are in the 1CO class.
For the 2CO class of bivalents - all the SC are longer, the high group averages were longer than the low.

 and the mean 2COs SC are higher in the high strains
(this must have something to do with the physical size effect)

# Q2 IFD

this is the Q2 prediction for IFD / interference

1. glm across subsp

2. anova within subsp

3. logistic regression for musc (or all males)

```{r Q2.IFD, echo=FALSE}

#glm

#anova for glm

#logistic regression for high and low group

> Male Polymorphism Prediction

test 3. logistic regression
$$  Rec \ Group ~=~ mouse \ average \ IFD $$
For these tests I am looking at the single IFD measure from 2CO bivalents.

```{r male.IFD_strain, echo=FALSE, eval=FALSE}
#integrate this into the setup scripts

## Normalized CO positions
